Attribute fusion in a latent process model for time series of graphs

### 2011-05-20 ssp2011.tex
Completed the proof of num triangle and scan statistics.

Fig.~\ref{fig:notional_change_point}.

@@

\\ \noindent
We are interested in
the problem of testing, for a $t \in \mathbb{N}$, the hypotheses that
$t$ is the change-point of $\mathscr{V}$, namely

@@

for $l \geq 2$. As we have mentioned earlier, we will use the first
approximation $\bar{\mathscr{V}}$ of $\mathscr{V}$ in our study of the 
asymptotic theory for these graphs invariants.   

@@

@@

  \frac{\tau_{\lambda}(G(t))}{\tbinom{n}{3}} &= \tbinom{n}{3}^{-1}

@@

  \\ &= \tbinom{n}{3}^{-1}
  \sum_{\mathclap{\{e_1, e_2, e_3\} \in \tbinom{E(t)}{3}}} Y_{e_1}
  Y_{e_2} Y_{e_3}
$\tau_{\lambda}(G(t))/\tbinom{n}{3}$ is then an (incomplete)

@@

  \tau_{\lambda}^{*}(G(t)) = \sum_{e \in E(t)} \mathbb{E}[
  \tau_{\lambda}(G(t)) \, \lvert \, Y_{e}]

@@

    \langle \pi_{01} \rangle^{2} \Bigr]^{2} \\ 

@@

@@

@@

@@

@@

\end{IEEEproof}
\bibliography{ssp2011}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

@@

Fig.~\ref{fig:notional_change_point}. We are interested in the problem
of testing, for a $t \in \mathbb{N}$, the hypotheses that $t$ is the
change-point of $\mathscr{V}$, namely

@@

for $l \geq 2$, i.e., we want to find a parameter $\lambda$ such that
the power of the test using $T_{\lambda}^{l}$ is maximum. We note that
$T_{\lambda}^{l}$ is scale invariant in $\lambda$ for all of our graph
invariants. It was noted in
\cite{lee:_laten_proces_model_time_attrib_random_graph} that the
maximum of the power for scale invariant tests is
attainable in $\{ \| \lambda \| \leq 1 \}$. \\ \\
As we have mentioned earlier, we will assume that the
number of vertices $n$ is large and that the use of the first order approximation
 $\bar{\mathscr{V}}$ of $\mathscr{V}$ is appropriate in our study of the 
asymptotic theory for these graphs invariants. We strive to
obtain estimates for the power in testing $t^{*} > t$ against $t^{*} =
t$ using $T_{\lambda}(t)$ and from these estimates, find the $\lambda$
that maximize the power. We now introduce a few
additional notations that will be used later on in the paper. First,
let $\pi_{00}, \pi_{01}$, and $\pi_{11}$ be sub-probability vectors
whose components are given by
  \pi_{00}(k) = \pi_{0}(k) \pi_{0}(k) \\
  \pi_{01}(k) = \pi_{0}(k) \pi_{1}(k) \\
  \pi_{11}(k) = \pi_{1}(k) \pi_{1}(k)
where $k \in \mathscr{K}$. We also let $\eta_{00}, \eta_{01}$, and
$\eta_{11}$ be matrices of size $K \times K$ defined by 
  \eta_{00} = \mathrm{diag}(\pi_{00}) - \pi_{00} \pi_{00}^{T} \\
  \eta_{01} = \mathrm{diag}(\pi_{01}) - \pi_{01} \pi_{01}^{T} \\
  \eta_{11} = \mathrm{diag}(\pi_{11}) - \pi_{11} \pi_{11}^{T}
% The structure of the remaining sections of the paper is as follows. We
% discussed the power estimates for $\mathcal{E}_{\lambda},
% \tau_{\lambda}, \Delta_{\lambda}$ and $\Psi_{\lambda}$ in
% \S~\ref{sec:power-estim-mathc}
%  through \S~\ref{sec:power-estim-psi_l}. We illustrate our analysis with some
% simulation experiments in \S~\ref{sec:experimental-results}  

@@

\label{sec:power-estim-mathc}

@@

  \frac{\tau_{\lambda}(G(t))}{\tbinom{n}{3}} = \tbinom{n}{3}^{-1}

@@

$\tau_{\lambda}(G(t))/\tbinom{n}{3}$ is then an

@@

  \tau_{\lambda}^{*}(G) - \mathbb{E}[\tau_{\lambda}(G)] =
  \sum_{\mathclap{\{u,v\} \in \tbinom{V(t)}{2}}} \, \Bigl(\mathbb{E}[
  \tau_{\lambda}(G) \, \lvert \, Y_{uv}] -
  \mathbb{E}[\tau_{\lambda}(G)]\Bigr)

@@

    \langle \lambda, \pi_{01} \rangle^{2} \Bigr]^{2} \\ 

@@

\label{sec:power-estim-delt}

@@

\label{sec:power-estim-psi_l}

@@

%\section{Experimental Results}
%\label{sec:experimental-results}

@@

  $\tau_{\lambda}(G)$ is a U-statistics with kernel function
  $h(Y_1, Y_2, Y_3) = Y_1 Y_2 Y_3$. By the theory of U-statistics, we
  know that
    \label{eq:48}
    \frac{\tau_{\lambda}(G) -
      \mathbb{E}[\tau_{\lambda}^{*}(G)]}{\sqrt{\mathrm{Var}[\tau_{\lambda}^{*}(G)]}}
    \rightsquigarrow N(0,1)
  provided that $\mathrm{Var}[\tau_{\lambda}(G) -
  \tau_{\lambda}^{*}(G)] = o(\mathrm{Var}[\tau_{\lambda}^{*}(G)]$. We
  will not show this here. The deta

By the independent edge
  assumption, we have 
    \mathbb{E}[h(Y_i, Y_j, Y_k) &= \mathbb{E}[Y_i]
  \mathbb{E}[Y_j] \mathbb{E}[Y_k] \\ 
  \mathbb{E}[h(Y_i, Y_j, Y_k) |
  Y_i] &= Y_i E[Y_j] \mathbb{E}[Y_k] 
 Thus, for $t < t^{*}$, we have $\mathbb{E}[\tau_{\lambda}^{*}(G(t))] = \tbinom{n}{3} \langle
  \lambda, \pi_{00} \rangle^{3}$ and
      \mathrm{Var}[\tau_{\lambda}^{*}(G(t))] &=
      \mathrm{Var}\Bigl[\sum_{\{u,v,w\}} Y_{uv} \mathbb{E}[Y_{uw}]
      \mathbb{E}[Y_{vw}]\Bigr] \\
      &= (n-2)^{2} \langle \lambda, \pi_{00} \rangle^{4}
      \mathrm{Var}\Bigl[\sum_{\{u,v\}} Y_{uv} \Bigr] \\
      &= (n-2)^{2} \langle \lambda, \pi_{00} \rangle^{4} \tbinom{n}{2}
      \langle \lambda, \eta_{00} \lambda \rangle
  The case when $t = t^{*}$ can be argued in a similar manner. We
  sketch here the derivation of
  $\mathrm{Var}[\tau_{\lambda}^{*}(G(t))]$. We can partition the set
  $\{u,v\} \in \tbinom{V}{2}$ into the sets 
\mathcal{S}_1 &= \{ u,v \in [m]
  \}, \\ \mathcal{S}_2 &= \{ u \in [m], v \in [n] \setminus [m]\}, \\
  \mathcal{S}_3 &= \{ u, v \in [n] \setminus[m]\}
\end{align*} Now, for $\{u,v\} \in \mathcal{S}_1$, we have
  \label{eq:58}
    \sqrt{S_1} &= \sum_{w \not \in \{u,v\}} \mathbb{E}[ h(Y_{uv},
    Y_{uw}, Y_{vw}) | Y_{uv}] \\ & = ((m-2) \langle \lambda, \pi_{11}
    \rangle^{2} + (n-m) \langle \lambda, \pi_{10} \rangle^{2})Y_{uv}
The above expression is reasoned as follows. If $w \in [m]$, then
$\mathbb{E}[Y_{uw}] = \mathbb{E}[Y_{vw}] = \langle \lambda,
\pi_{11}$ and there are $m-2$ possible choices for $w \in [m]$
different from $u$ and $v$. If $w \in [n] \setminus [m]$, then
$\mathbb{E}[Y_{vw}] = \mathbb{E}[Y_{uw}] = \langle \lambda, \pi_{10}
\rangle$ and there are $n - m$ possible choices for $w$. Analogous
reasoning gives the expressions for $S_2$ and $S_3$ in the statement
of the lemma. We therefore have
  \label{eq:60}
  \mathrm{Var}[\tau_{\lambda}^{*}(G(t))] &=  \mathrm{Var}\Bigl[\sum_{\{u,v,w\}} Y_{uv} \mathbb{E}[Y_{uw}]
      \mathbb{E}[Y_{vw}]\Bigr] \\ &=
      S_1 \mathrm{Var}[\sum_{\{u,v\} \in \mathcal{S}_1} Y_{uv}] + S_2
      \mathrm{Var}[\sum_{\{u,v\} \in \mathcal{S}_1} Y_{uv}] \\ &+ S_3
      \mathrm{Var}[\sum_{\{u,v\} \in \mathcal{S}_1} Y_{uv}] \\ &=
      \tbinom{m}{2} \langle \lambda, \eta_{00} \lambda \rangle S_1 +
      m(n-m) \langle \lambda, \eta_{01} \lambda \rangle S_2 \\ &+
      \tbinom{n-m}{2} \langle \lambda, \eta_{00} \lambda \rangle S_3
as desired. To complete the proof one must show that
$\mathrm{Var}[\tau_{\lambda}(G) - \tau_{\lambda}^{*}(G)] =
o(\mathrm{Var}[\tau_{\lambda}^{*}(G)]$ and this follows directly from
the corresponding argument in \cite{nowicki88:_subgr_u_statis_method}
or \cite{rukhin09:_asymp_analy_various_statis_random_graph_infer}.

@@

From Eq.~\eqref{eq:27} and Eq.~\eqref{eq:30}, we have
  \label{eq:40}
 \lim
\mathbb{P}(\Upsilon(G) \geq a_{n,m}) = \lim \mathbb{P}(\Delta(G) \geq
N_{\kappa})
Let $N_{\kappa,y} = N_\kappa + y \tfrac{\sigma_{E+F}}{\sqrt{2
    \log{m}}}$. We now define $a_{n,m,y}$ as 
  \label{eq:18}
  \begin{split}a_{n,m,y} &= \langle \lambda, \pi_{00} \rangle
    \tbinom{N_{\kappa,y}}{2} +
  \langle \lambda, \pi_{11} - \pi_{00} \rangle \tbinom{\mu_E}{2}  +
  \langle \lambda, \pi_{10} - \pi_{00} \rangle \mu_E \mu_F \\
  &= a_{n,m} + \langle \lambda, \pi_{00} \rangle y
  \frac{\sigma_{E+F}}{\sqrt{2 \log m}} \Bigl(N_{\kappa} + y
  \frac{\sigma_{E+F}^{2}}{2 \sqrt{2 \log{m}}} + O(1) \Bigr) \\
  &= a_{n,m} + (y + o(1)) b_{n,m} 
We therefore have
  \label{eq:43}
  \lim \mathbb{P}(\Upsilon(G) \geq a_{n,m,y}) &= 
  \lim \mathbb{P}\Bigl(\frac{\Upsilon(G) - a_{n,m}}{b_{n,m}} \geq y \Bigr) \\
  &= \lim \mathbb{P}(\Delta(G) \geq N_{\kappa,y}) \\
  &= \lim \mathbb{P}\Bigl(\frac{\Delta(G) -
    N_{\kappa}}{\sigma_{E+F}} \geq \frac{y}{\sqrt{2 \log{m}}} \Bigr) \\
Because $\Delta(G)$ converges weakly to a Gumbel
distribution in the limit
(\cite{bollobas85:_random_graph,rukhin:_limit_distr_graph_scan_statis}),
  \label{eq:47}
  \mathbb{P}\Big(\frac{\Upsilon(G) - a_{n,m}}{b_{n,m}} \leq
    y\Bigr) \rightarrow e^{- e^{-y}} 
and Eq.~\eqref{eq:29} follows.
\bibliography{ssp2011}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:
